---
title: "Is Management Trying to Cover up the Performance of Their Company?"
subtitle: "Unveiling Red Flags in Management's Discussion and Analysis"
author: 
  - Nikhil Iyer
thanks: "Code and data are available at: https://github.com/Niyer02/sec-market-comparison"
date: today
date-format: long
abstract: "Every quarter, management must release a document outlining the performance of their company, however, these documents often reveal nothing to the naked eye. Through lengthy documents, abundant filler words, and a lack of transparency, understanding these red flags is crucial for stakeholders to make informed decisions."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

```


# Introduction

While blatant attempts to conceal aspects of their company have often become headlines, the subtler attempts often go
under the radar. More filler words and longer sections have become a staple in almost every SEC filing. Understanding these subtle flags
is paramount for stakeholders to safeguard their portfolios.

This paper aims to understand SEC filings, and demonstrate the correlation between the Management Discussion and Analysis (MD&A) section, and the value of the stock. Using R [@citeR], we are able to process the retrieved data, and build a model that accurately represents the trend between MD&A's and the price of the stock.


# Data {#sec-data}


```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

```


```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

```


# Model


## Model set-up

### Model justification



# Results


```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| warning: false
```




# Discussion

## First discussion point {#sec-first-point}


## Second discussion point

## Third discussion point

## Weaknesses and next steps


\newpage

\appendix

# Appendix {-}


# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
```

## Diagnostics


```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| layout-ncol: 2

```



\newpage


# References


